### 2014-10-27 intro.tex
intro and motive

Many real-world web services --- such as those built and maintained by Amazon,
Facebook, Google, Twitter, etc. --- replicate application state and logic
across multiple \emph{replicas} within and across data centers. Replication is
intended not only to improve application throughput and reduce user-perceived
latency, but also to tolerate partial failures without compromising overall
service availability. Traditionally programmers have relied on \emph{strong
consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct applications.
While strong consistency is an easily stated property, it masks the reality
underlying large-scale distributed systems with respect to non-uniform latency
(i.e., availability) and network partitions~\cite{Brewer2000,Gilbert}. Indeed,
modern web services, which aim to provide an "always on" experience,
overwhelmingly favour availability and partition tolerance over strong
consistency. To this end, several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have consequently been proposed.

Under weak consistency, the developer needs to be aware of concurrent
conflicting updates, and has to pay careful attention to avoid unwanted
inconsistencies (e.g., negative balances in a bank account, or having an item
appear in a shopping cart after it has been removed~\cite{Dynamo}). Oftentimes,
the inconsistency leaks from the application and is witnessed by the user.
Ultimately, the developer must decide the consistency level appropriate for a
particular operation; this is understandably an error-prone process requiring
intricate knowledge of both the application as well as the semantics and
implementation of the underlying data store.  More often than not, only
informal descriptions are available~\cite{}.  Picking the correct consistency
level is critical not only for correctness but also for scalability of the
application. While choosing a weaker consistency level than required may
introduce program errors and anomalies, choosing a stronger one than necessary
level can negatively impact program scalability.

@@

Moreover, weak consistency hinders compositional reasoning about programs.
While an application might be naturally expressed in terms of well-understood
and expressive data types such as maps, trees, queues, or graphs,
geo-distributed stores typically only provide a minimal set of data types with
in-built conflict resolution strategies such as last-writer-wins (LWW)
registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermor, while
traditional database systems enable composability through transactions,
existing geo-distributed stores lack unrestricted transactional access to the
data.  Working with geo-distributed stores thus requires application state to
be suitably coerced to work within the capabilities of the store.

To address these issues, we describe \name, a declarative programming model and
implementation for eventually consistent stores. The key novelty of \name is an
expressive \emph{contract language} to declare and verify fine-grained
application-level consistency properties. The programmer uses the contract
language to axiomatically describe the set of legal executions allowed over the
replicated data type. Contracts are constructed using primitive consistency
relations such as \emph{visibility} and \emph{session order} along with
standard logical and relational operators. A \emph{contract enforcement system}
automatically maps each operation to a particular consistency level available
on the store. In addition, the system statically reports an error if the given
application-level consistency property cannot be satisfied by the store.

The paper makes the following contributions:

\begin{itemize}

\item We introduce \name, a language for declaratively describing eventually
	consistent programs that manipulate replicated data types. Contracts are used
	to specify fine-grained application-level consistency properties, and are
	analyzed to assign the most efficient and sound store consistency level to
	the corresponding operation.

\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct and most efficient isolation level for a transaction.

\item Metatheory that certifies the soundness of our contract enforcement
	system, and ensures that an operation is only executed if the required
	conditions on consistency are met is also given.

\item \name is implemented as a shim layer over (and importantly, without
	modifying) Cassandra~\cite{Cassandra}, a well-known general-purpose data
	store.  An extensive experimental evaluation over a collection of real-world
	applications, including a Twitter-like micro-blogging site and an eBay-like
	auction site illustrates the practicality of our approach.

\end{itemize}

>>>>

Many real-world web services --- such as those built and maintained by
Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
logic across multiple \emph{replicas} within and across data
centers. Replication is intended not only to improve application throughput
and reduce user-perceived latency, but also to tolerate partial failures
without compromising overall service availability. Traditionally programmers
have relied on \emph{strong consistency} guarantees such as
linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct
applications.  While strong consistency is an easily stated property, it
masks the reality underlying large-scale distributed systems with respect to
non-uniform latency (i.e., availability) and network
partitions~\cite{Brewer2000,Gilbert}. Indeed, modern web services, which aim
to provide an "always on" experience, overwhelmingly favour availability and
partition tolerance over strong consistency. To this end, several \emph{weak
  consistency} models such as eventual consistency, causal consistency,
session guarantees, and timeline consistency have been proposed.

Under weak consistency, the developer needs to be aware of concurrent
conflicting updates, and has to pay careful attention to avoid unwanted
inconsistencies (e.g., negative balances in a bank account, or having an
item appear in a shopping cart after it has been
removed~\cite{Dynamo}). Oftentimes, the inconsistency leaks from the
application and is witnessed by the user.  Ultimately, the developer must
decide the consistency level appropriate for a particular operation; this is
understandably an error-prone process requiring intricate knowledge of both
the application as well as the semantics and implementation of the
underlying data store, which typically have only informal
descriptions~\cite{}.  Nonetheless, picking the correct consistency level is
critical not only for correctness but also for scalability of the
application. While choosing a weaker consistency level than required may
introduce program errors and anomalies, choosing a stronger one than
necessary can negatively impact program scalability.

@@

Weak consistency also hinders compositional reasoning about programs.  While
an application might be naturally expressed in terms of well-understood and
expressive data types such as maps, trees, queues, or graphs,
geo-distributed stores typically only provide a minimal set of data types
with in-built conflict resolution strategies such as last-writer-wins (LWW)
registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermore, while
traditional database systems enable composability through transactions,
geo-distributed stores typically lack unrestricted transactional access to
the data.  Working in this environment thus requires application state to be
suitably coerced to function using only the capabilities of the store.

To address these issues, we describe \name, a declarative programming model
and implementation for eventually consistent geo-distributed data
stores. The key novelty of \name is an expressive \emph{contract language}
to declare and verify fine-grained application-level consistency
properties. The programmer uses the contract language to axiomatically
specify the set of legal executions allowed over the replicated data
type. Contracts are constructed using primitive consistency relations such
as \emph{visibility} and \emph{session order} along with standard logical
and relational operators. A \emph{contract enforcement system} automatically
maps operations over the datatype to a particular consistency level
available on the store, and provably validates the correctness of the
mapping.  The paper makes the following contributions:

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a language for declaratively describing eventually
	consistent programs that manipulate replicated data types. Contracts are used
	to specify fine-grained application-level consistency properties, and are
	analyzed to assign the most efficient and sound store consistency level to
	the corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct and most efficient isolation level for a transaction.
\item We provide metatheory that certifies the soundness of our contract enforcement
	system, and ensures that an operation is only executed if the required
	conditions on consistency are met.
\item \name is implemented as a transparent shim layer over
  Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
  extensive experimental evaluation over a collection of real-world
  applications, including a Twitter-like micro-blogging site and an
  eBay-like auction site illustrates the practicality of our approach.
\end{itemize}

### 2014-10-27 motivation.tex
intro and motive

In this section, we motivate the need and benefit of declarative reasoning under
eventual consistency with the help of an example. We begin with a description of
our system and programming model, and introduce the primitive relations that
constitute an essential part of our contract language.

@@

Each object is associated with a set of \emph{operations}. The clients interact
with the store by invoking the operations on objects. The sequence of operations
invoked by a particular client on the store is called a \emph{session}. The
data store is typically accessed by a large number of clients (and hence
sessions) concurrently. Importantly, the clients are oblivious to which replica
an operation is applied to; the data store may choose to route the operation to
the closest replica for minimizing latency, or route it to a different replica
in order to balance the load in the system. For example, the operations $foo$
and $bar$ invoked by the same session on the same object, might end up being
applied to different replicas because replica 1 to which $foo$ was applied to
might be unreachable when invoking $bar$.

Suppose an operation $foo$ is invoked on a object $x$ with arguments $arg_1$ at
replica 1. The operation $foo$ simply \emph{reduces} over the current set of
effects at that replica on that object ($w^x_1$ and $w^x_2$), produces a result
$v1$ that is sent back to the client, and emits a \emph{single} new effect
$w^x_4$, which is appended to the state $x$ at replica 1. Thus, every operation
is evaluated over a \emph{snapshot} of the state of the object on which it is
invoked. In this case, we say that the effect $w^x_4$ \emph{witnessed} $w^x_1$
and $w^x_2$. Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
$\visZ$ is the visibility relation between the effects. Visibility is an
irreflexive (an operation cannot see its own effect) and asymmetric relation,
and only relates effects produced by operations on the same object.

Read-only operation execution is similar except that they do not produce a new
effect. The effect added to a particular replica is asynchronously sent to
other replicas. We assume that \emph{eventually}, an effect applied at one
replica will be merged into all other replicas. The two effects $w^x_4$ and
$w^x_5$ that arise from the same session are said to be in \emph{session order}
(written logically as $\so{w^x_4}{w^x_5}$). Session order is a irreflexive,
transitive relation. The effects $w^x_4$ and $w^x_5$ arising from operations
applied to the same object $x$ are said to be under the \emph{same object}
relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every
effect with the operation that generated the effect with the help of a relation
$\operZ$. In the current example, $\oper{w^x_4}{foo}$ and $\oper{w^x_5}{bar}$
hold. For simplicity, we assume that all the operation names across all object
types are distinct.

The system model described allows all the inconsistencies associated with
eventual consistency. The goal of our work is to identify the precise
consistency level for each operation such that the application-level
constraints are not violated. In the next section, we will concretely describe
the challenges associated with constructing a consistent bank account on top of
eventually consistent data stores. Subsequently, we will illustrate how our
contract language, armed with the primitive relations $\visZ$, $\soZ$,
$\sameobjZ$ and $\operZ$, mitigates these challenges.

@@

Suppose our goal is to implement a highly available bank account on top of an
eventually consistent data store, with the \emph{integrity} constraint that the
balance should never drop below 0. Let us assume that the bank account object
provides three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance},
with the assumption that the withdraw fails if the account has insufficient
balance to complete the withdraw operation. Our goal is to choose the correct
consistency level for each operation such that (1) the balance never drops
below zero and (2) the \cf{getBalance} operation never incorrectly show a
negative balance. To be able to do this, we need to consider the anomalies that
could arise under eventual consistency.

@@

Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
that all operations in the figure are on the same bank account object with
initial balance as 0. Session 1 performs a \cf{deposit} of 100, followed by a
\cf{withdraw} of 80 in the same session. The \cf{withdraw} operation witnesses
the deposit and succeeds\footnote{Although visibility and session order
relations relate effects, we have abused the notation in these examples to
relate operations, with the idea that the relations relate the effect emitted
by those operations}. Subsequently, session 2 perform a \cf{withdraw}
operation, but importantly, due to eventual consistency, only witnesses the
\cf{deposit} from session 1, but not the subsequent withdraw. Hence, this
\cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
constraint. A subsequent \cf{getBalance} operation, that witnesses all the
previous operations and reports a negative balance.

It is easy to see that the only way to ensure that integrity constraint on
\cf{balance} is to prevent concurrent withdrawals on the same bank account
object. This can be done by insisting that \cf{withdraw} be executed with
strong consistency. Although this prevents the integrity constraint from being
violated, a strongly consistent \cf{withdraw} itself is not sufficient to
prevent \cf{getBalance} from ever reporting a negative balance to the end user.

Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly}, which
consists of three concurrent sessions performing a \cf{deposit}, a
\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
account object. Assume that initial balance in the account is zero. As the
\cf{vis} edge indicates, operation \cf{withdraw(50)} in session 2, can
witnesses the effects of \cf{deposit(100)} from session 1, leading it to
rightfully conclude that the current balance is 100. Consequently,
\cf{withdraw(50)} is performed safely. The effect of this \cf{withdraw}
operation subsequently becomes visible to the \cf{getBalance} in Session 3.
However, \cf{getBalance}, which only witnesses the effect of \cf{withdraw} from
session 2, but not the \cf{deposit} from session 1, reports the balance of
negative 50 to the user.

Negative balance is not the only inconsistency that the user is exposed to
under an eventually consistent bank account. Figure
~\ref{fig:missingUpdateAnomaly} shows an execution where the \cf{getBalance}
operation in a session does not witness the effects of \cf{withdraw} operation
performed in the same session, possibly because \cf{getBalance} was served by a
replica that has not yet merged the \cf{withdraw} effect, served by a different
replica. This anomaly leads the user to incorrectly conclude that the
\cf{withdraw} operation failed to go through.

Although it is easy to understand the reasons behind the occurrence of the
aforementioned negative balance and missing update anomalies, the fix is not
quite apparent. Making \cf{getBalance} a strongly consistent operation is
definitely sufficient to avert anomalies, but is it really necessary? Given
that a strongly consistent operation is significantly expensive, both in terms
of availability and monetary overheads~\cite{DynamoDB,Pileus}, we ideally would
not want to request strong consistency, unless there are no alternatives
available. Unfortunately, this requires us to understand subtle differences in
semantics of various kinds of weak consistency alternatives, offered by the
data stores.

@@

\name automates this error-prone task by letting the programmer declare
application-level consistency constraints as \emph{contracts}. The key
intuition is that, in the case of the bank account example, any execution that
does not exhibit the anomalies described in the previous section is a
\emph{well-formed} execution on the bank account object. In the same vein as
axiomatic memory models for shared memory multicore
processors~\cite{Demange2013,Sarkar2011}, the programmer axiomatically
describes the set of legal executions on top of each replicated date type over
a trace of operation invocations on that type. Quelea ensures that any
execution on top of such datatypes always belongs to the set of well-formed
executions on that type.

In the bank account example, we observed that, in order to preserve the
integrity constraint, the \cf{withdraw} operation must be strongly consistent.
That is, given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible
to $b$ or vice versa. Utilizing the primitive relations described in
Section~\ref{sec:sysmod}, we express this application-level consistency
requirement as a contract over \cf{withdraw}:
\begin{smathpar}
\begin{array}{l}
\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
\end{array}
\end{smathpar}
In the above contract, $\cureff$ stands for the effect emitted by the
\cf{withdraw} operation. The syntax $a : \rcf{withdraw}$ states that $a$ is an
effect which was emitted by a \cf{withdraw} operation i.e)
$\oper{a}{\rcf{withdraw}}$ holds. The contract says that, if the current
operation emits an effect $\cureff$, then for any operation $a$ which was
emitted by a \cf{withdraw} operation, it is the case that $a = \cureff$ or $a$
is visible to $\cureff$ or vice versa. Any execution on a bank account object
that preserves the above contract for a \cf{withdraw} operation is said to have
a correct implementation of \cf{withdraw}.

For the \cf{getBalance}, we construct a contract  precisely to prevent the
anomalies described previously. The contract is shown below:
\begin{smathpar}
\begin{array}{l}
\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
\end{array}
\end{smathpar}

If a \cf{withdraw} $b$ is visible to the get balance operation, then all
\cf{deposit} operations $a$ visible to the \cf{withdraw} should be visible to
the get balance operation. This prevents negative balance anomalies. The syntax
$c : \rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an effect that was
emitted either by a \cf{deposit} operation or by a \cf{withdraw} operation. The
contract language provides relational operators to compose relations. The
syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The
last line of the above contract says that if a \cf{deposit} or a \cf{withdraw}
precedes a \cf{getBalance} operation in session order, and is applied on the
same object as the \cf{getBalance} operation, then it must be the case that the
\cf{getBalance} operation witnesses the effects of the preceding operations.
Since the deposit operation does not have any restrictions, its contract is
simply $\true$.

\subsection{From contracts to implementation}

Notice that the contracts for \cf{withdraw} and \cf{getBalance} only express
the application-level consistency requirements, and make no reference to the
semantics of the underlying store. To write such contracts, a programmer only
needs to reason about the semantics of the application under the \name system
model. The mapping of application-level consistency requirements to appropriate
store-level guarantees is done automatically behind the screen.

How might one go about ensuring that any execution adheres to the contract? The
challenge is that contract is a declarative characterization of correctness of
execution, while what we need is an operational procedure for \emph{enforcing}
the contracts. One strategy is to adopt speculative execution as espoused by
other eventually consistent systems, where operations are tentatively applied
as they are received from the client or other replicas. We can maintain a
runtime manifestation of the axiomatic characterization and check the
well-formedness at runtime, and rollback the execution if
ill-formed~\cite{rxcml}. However, the overhead of state maintenance and the
complexity of user-defined contracts (unlike simple tests such as cyclic
dependence) make this technique unviable.

Instead, Quelea statically analyzes the contracts using an SMT solver, and maps
each operation to a particular store-level consistency under which the contract
would be preserved. In particular, the key idea is that the system implementor
axiomatically describes the store semantics in the same contract language as
the one used to express the contract on the operation. We translate the
contracts down to decidable subset of first-order logic~\cite{}, and discharge
proof obligation of checking whether a particular store consistency level is
sufficient for the given contract with the help of an SMT solver. The next
section describes the contract language and classification.

>>>>

In this section, we motivate the need and benefit of declarative reasoning
for eventually consistent systems.  We begin with a description of our
system and programming model, and introduce the primitive relations that
constitute the essential part of our contract language.

@@

Each object is associated with a set of \emph{operations}. The clients
interact with the store by invoking operations on objects. The sequence of
operations invoked by a particular client on the store is called a
\emph{session}. The data store is typically accessed by a large number of
clients (and hence sessions) concurrently. Importantly, the clients are
oblivious to which replica an operation is applied to; the data store may
choose to route the operation to the closest replica for minimizing latency,
or route it to a different replica in order to balance the load in the
system. For example, the operations \emph{foo} and \emph{bar} invoked by the same
session on the same object, might end up being applied to different replicas
because replica 1 (to which \emph{foo} was applied) might be unreachable when
the client invokes \emph{bar}.

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of
effects at that replica on that object ($w^x_1$ and $w^x_2$), produces a result
$v1$ that is sent back to the client, and emits a \emph{single} new effect
$w^x_4$ that is appended to the state $x$ at replica 1. Thus, every operation
is evaluated over a \emph{snapshot} of the state of the object on which it is
invoked. In this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$
and $w^x_2$. Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
$\visZ$ is the visibility relation between effects. Visibility is an
irreflexive (an operation cannot see its own effect) and asymmetric relation,
and only relates effects produced by operations on the same object.
Executing a read-only operation is similar except
that no new effects are produced.

The effect added to a particular replica is asynchronously sent to other
replicas. We assume that \emph{eventually} an effect applied at one replica
will be merged into all other replicas. Two effects $w^x_4$ and $w^x_5$ that
arise from the same session are said to be in \emph{session order} (written
logically as $\so{w^x_4}{w^x_5}$). Session order is an irreflexive,
transitive relation. The effects $w^x_4$ and $w^x_5$ arising from operations
applied to the same object $x$ are said to be under the \emph{same object}
relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every
effect with the operation that generated the effect with the help of a
relation $\operZ$. In the current example, $\oper{w^x_4}{foo}$ and
$\oper{w^x_5}{bar}$ hold. For simplicity, we assume all operation names
across all object types are distinct.

This model admits all the inconsistencies associated with eventual
consistency. The goal of this work is to identify the precise consistency
level for each operation such that application-level constraints are not
violated. In the next section, we will concretely describe the challenges
associated with constructing a consistent bank account on top of an
eventually consistent data stores. Subsequently, we will illustrate how our
contract and specification language, armed with the primitive relations
$\visZ$, $\soZ$, $\sameobjZ$ and $\operZ$, mitigates these challenges.

@@

Consider how we might implement a highly available bank account on top of an
eventually consistent data store, with the \emph{integrity} constraint that
the balance should never drop below 0. Let us assume that the bank account
object provides three operations: \cf{deposit}, \cf{withdraw} and
\cf{getBalance}, with the assumption that the withdraw fails if the account
has insufficient balance. Our goal is to choose the correct consistency
level for each of these operations such that (1) the balance never drops
below zero and (2) the \cf{getBalance} operation never incorrectly show a
negative balance. To be able to do this, we need to consider the anomalies
that could arise under eventual consistency.

@@

Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
that all operations in the figure are on the same bank account object with
the initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed by a
\cf{withdraw} of 80 in the same session. The \cf{withdraw} operation witnesses
the deposit and succeeds\footnote{Although visibility and session order
relations relate effects, we have abused the notation in these examples to
relate operations, with the idea that the relations relate the effect emitted
by those operations}. Subsequently, session 2 perform a \cf{withdraw}
operation, but importantly, due to eventual consistency, only witnesses the
\cf{deposit} from session 1, but not the subsequent withdraw. Hence, this
\cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
constraint. A subsequent \cf{getBalance} operation, that happens to witness all the
previous operations,  would report a negative balance.

It is easy to see that the only way to ensure the integrity constraint on
\cf{balance} is to prevent concurrent withdrawals on the same bank account
object. This can be done by insisting that \cf{withdraw} be executed with
strong consistency properties. Although this prevents the integrity
constraint from being violated, a strongly consistent \cf{withdraw} itself
is not sufficient to prevent \cf{getBalance} from ever reporting a negative
balance to the end user.

Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly}, which
consists of three concurrent sessions performing a \cf{deposit}, a
\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same
bank account object. As before, assume that the initial balance in the
account is zero. As the \cf{vis} edge indicates, operation \cf{withdraw(50)}
in session 2, can witnesses the effects of \cf{deposit(100)} from session 1,
leading it to rightfully conclude that the current balance is
50. Consequently, \cf{withdraw(50)} is performed safely. The effect of this
\cf{withdraw} operation subsequently becomes visible to the \cf{getBalance}
in Session 3.  However, because \cf{getBalance} only witnesses the effect of
the \cf{withdraw} operation from session 2, but not the \cf{deposit} from
session 1, reports the balance of negative 50 to the user.

Producing a negative balance is not the only inconsistency that the user is
exposed to under an eventually consistent bank account. Figure
~\ref{fig:missingUpdateAnomaly} shows an execution where the \cf{getBalance}
operation in a session does not witness the effects of an earlier
\cf{withdraw} operation performed in the same session, possibly because it
was served by a replica that has not yet merged the \cf{withdraw} effect,
served by a different replica. This anomaly leads the user to incorrectly
conclude that the \cf{withdraw} operation failed to go through.

%% \begin{figure}
%% \centering
%% \includegraphics[width=\columnwidth]{Graphs/BankAccount_Th_vs_Lat}
%% \caption{Bank Account: Throughput vs Latency. Quelea is our implementation.
%% Strong and Quelea are correct. High. Avail. and causal are incorrect (withdraw
%% not strong). 20\% withdraw, 20\% deposit, 60\% getBalance.}
%% \label{grf:BankAcc_Th_vs_Lat}
%% \end{figure}

Although it is easy to understand the reasons behind the occurrence of the
aforementioned negative balance and missing update anomalies, finding the
appropriate fixes is not readily apparent. Making \cf{getBalance} a strongly
consistent operation is definitely sufficient to avert anomalies, but is it
really necessary? Given the cost of enforcing strong
consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid unless there
are no viable alternatives.  Exploring the space of these alternatives
requires understanding the subtle differences in semantics of various kinds
of weak consistency alternatives.

%%  a strongly consistent operation is
%% significantly expensive (Figure~\ref{grf:BankAcc_Th_vs_Lat} \KC{Should I
%%   include the numbers here? If yes, I'll expand the text.}), both in terms
%% of availability and monetary overheads~\cite{DynamoDB,Pileus}, we ideally
%% would not want to request strong consistency, unless there are no
%% alternatives available.

@@

\name helps facilitate the mapping of operations to appropriate consistency
levels by letting the programmer declare application-level consistency
constraints as \emph{contracts} that axiomatically specify the set of
allowed executions involving this operation.  In the case of the bank
account example, any execution that does not exhibit the anomalies described
in the previous section is a \emph{well-formed} execution on the bank
account object.  By specifying the set of legal executions for each data
type in terms of a trace of operation invocations on that type,
\name\ ensures that all executions over that type are well-formed.

In our running example, it is clear that in order to preserve the integrity
constraint, the \cf{withdraw} operation must be strongly consistent.  That
is, given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to
$b$ or vice versa. Utilizing the primitive relations described in
Section~\ref{sec:sysmod}, we express this application-level consistency
requirement as a contract or specification over \cf{withdraw}:
\begin{smathpar}
\begin{array}{l}
\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
\end{array}
\end{smathpar}
Here, $\cureff$ stands for the effect emitted by the \cf{withdraw}
operation. The syntax $a : \rcf{withdraw}$ states that $a$ is an effect
which was emitted by a \cf{withdraw} operation i.e.,
$\oper{a}{\rcf{withdraw}}$ holds.  The contract specifies that if the
current operation emits an effect $\cureff$, then for any operation $a$
which was emitted by a \cf{withdraw} operation, it is the case that $a =
\cureff$ or $a$ is visible to $\cureff$, or vice versa. Any execution on a
bank account object that preserves the above contract for a \cf{withdraw}
operation is said to be derived from a correct implementation of
\cf{withdraw}.

For \cf{getBalance}, we construct the following contract:
\begin{smathpar}
\begin{array}{l}
\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
\end{array}
\end{smathpar}

If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
operations $a$ visible to the \cf{withdraw} should also be visible to
\cf{getBalance}. This prevents negative balance anomalies. The expression $c
: \rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an effect that was
emitted either by a \cf{deposit} or a \cf{withdraw} operation.  Additonally,
our specification language provides relational operators to compose
relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b)
\wedge R_2(a,b)$. The last line of the above contract says that if a
\cf{deposit} or a \cf{withdraw} precedes a \cf{getBalance} operation in
session order, and is applied on the same object as the \cf{getBalance}
operation, then it must be the case that the \cf{getBalance} operation
witnesses the effects of the preceding operations.  Since there are no
restrictions on when and how a \cf{deposit} operation can execute, its
contract is simply $\true$.

\subsection{From Contracts to Implementation}

Notice that the contracts for \cf{withdraw} and \cf{getBalance} only express
application-level consistency requirements, and make no reference to the
semantics of the underlying store. To write contracts, a programmer only
needs to reason about the semantics of the application under the \name
system model. The mapping of application-level consistency requirements to
appropriate store-level guarantees is done automatically behind-the-scene.

How might one go about ensuring that any execution adheres to the contract?
The challenge is that a contract provides a declarative (axiomatic)
specification of an execution, while what is required is is an operational
procedure for \emph{enforcing} its implicit constraints. One strategy would
be to execute operations speculatively.  Here, operations are tentatively
applied as they are received from the client or other replicas. We can
maintain a runtime manifestation of executions, and check well-formedness
conditions at runtime, rolling back executions if they are
ill-formed. However, the overhead of state maintenance and the complexity of
user-defined contracts make this technique infeasible in practice.

We devise a static approach instead.  Contracts are analyzed with the help
of a theorem prover, and statically mapped to a particular store-level
consistency property that the prover guarantees preserves contract
semantics. The key idea is that the system implementor axiomatically
describes the store semantics in the \emph{same} contract language as the
one used to express the contract on the operation.
